Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-08-22 doc/response_to_reviewer_R1.Rmd

•	The methodology should clearly state whether the 30 species selected for the study represent the functional groups. It should include the number of species per Deciduousness group.

RESPONSE:

•	I think it is important to add the average of non-dry years in Figure 1a (without 2010, 2015 and 2020). This helps clarify the difference between a normal/rainy year and a dry year.

**RESPONSE: Thanks for this comment. We have now included the long-term mean trend along with the long-term standard deviation to for Fig 1a to show the deviation more clearly. Fig 1b now follows more closely, showing the deviation in SD units from the long-term mean.**

@@

162-163: Until when was the measurement? It is important to add the last year of measurement.

RESPONSE:

239-242: I think that this extra information is unnecessary, since although this measurement is interesting, it was not the one applied in this study.

**RESPONSE: We have now removed this statement regarding the alternate measure of deciduousness that was considered.**

@@

Table 1: “other size effect” is DBH?

RESPONSE:

Figure 1a: “Dy” is not correct

**RESPONSE: Thanks for noticing this error. We have fixed this to say "Day of year".**

@@

3 In the case of deciduousness, the authors indicated that they used the deciduous index from Williams et al. (2008). Since different species might shed leaves in different season. Did the authors consider the impact of the timing of deciduousness on the growth?

RESPONSE:

Fig. S1 What does the shade area indicate? The three different drought events?

**RESPONSE: Yes, the shaded area indicates the three different drought years. We have made edited the text to reflect this.**

@@

Fig S6 The R Square is too low, so it is difficult to detect the linear relationship. The same issue for Fig. S11 and Fig. S12. For the non-significant correlation, it’s better use dashed lines instead of solid line.

RESPONSE:

Fig S7 “…included in the analyses in the 2010 and 2015 datasets”, do you mean “2010, 2015 and 2020”?

**RESPONSE: Thanks for notiicing this error. We have changed this to "2010, 2015 and 2020".**

### 2025-08-22 doc/manuscript_hkk_drought_sensitivity.Rmd
Few comments addressed

We assigned deciduousness as a species-level trait using data from @williams_deciduousness_2008. 
We defined deciduousness as the mean over trees of the proportion of crown leaf loss at maximum, with a score ranging from 0 to 4 where 0 is most evergreen and 4 is most deciduous. 
For each species, these were calculated from observations of individual trees that were assigned categorical scores, and averaged, allowing species to take continuous, decimal values in the 0-4 range.
Data were unavailable for _Alphonsea ventricosa_, an evergreen species, and so we manually assigned it a deciduousness score of 1. 
We also considered an alternative measure of deciduousness - the duration of time spent at maximum crown loss.
This was strongly correlated with the proportion of crown loss; we chose to use the proportion of crown loss because it was available for more species.  
<!-- At the species level, we also calculated the maximum size across all individuals in the plot as measured in the 2008 whole-plot census.
We calculated niche breadth as the standard deviation of TWI across all observed individuals of each species in the whole plot census in 2008.  -->

>>>>

We assigned deciduousness as a species-level trait using data from @williams_deciduousness_2008. 
We defined deciduousness as the mean over trees of the proportion of crown leaf loss at maximum, with a score ranging from 0 to 4 where 0 is most evergreen and 4 is most deciduous. 
For each species, these were calculated from observations of individual trees that were assigned categorical scores, and averaged, allowing species to take continuous, decimal values in the 0-4 range.
Data were unavailable for _Alphonsea ventricosa_, an evergreen species, and so we manually assigned it a deciduousness score of 1. 
<!-- We also considered an alternative measure of deciduousness - the duration of time spent at maximum crown loss. -->
<!-- This was strongly correlated with the proportion of crown loss; we chose to use the proportion of crown loss because it was available for more species.   -->
<!-- At the species level, we also calculated the maximum size across all individuals in the plot as measured in the 2008 whole-plot census.
We calculated niche breadth as the standard deviation of TWI across all observed individuals of each species in the whole plot census in 2008.  -->

### 2025-08-22 doc/supplementary_information.Rmd
updated SI file with one more figure
Few comments addressed

![**Figure S1:** Raw climate data timeseries from 2009 to 2023 from ERA5Land and CHIRPS. Measurements represent daily rolling means over a 30-day time window for precipitation, max temperature and VPD and rolling sum for number of dry days.](display/climvars_fullseries_remote.png)

![**Figure S2:** Climate anomalies for all variables from 2009 to 2020. Values represent number of standard deviations from the long-term mean for daily rolling mean values.](display/climvars_anomalies_fullseries_remote.png)

@@

![**Figure S4:** Species sensitivities in drought years 2010 and 2015. Boxplots represent 25th, 50th (median) and 75th percentile of sensitivity for each species in each year and whiskers represent 95% CIs. All individual sensitivities represented by jittered points.](display/Fig_SI_species_sensitivities.png)

@@

![**Figure S7:** Map of trees with dendrobands in HKK, included in analyses in the 2010 and 2015 datasets](display/dendro_tree_map.png)

@@

>>>>

![**Figure S1:** Raw climate data timeseries from 2009 to 2023 from ERA5Land and CHIRPS. Measurements represent daily rolling means over a 30-day time window for precipitation, max temperature and VPD and rolling sum for number of dry days. The shaded areas indicate the three drought years analysed - 2010, 2015 and 2020.](display/climvars_fullseries_remote.png)

![**Figure S2:** Climate anomalies for all variables from 2009 to 2020. Values represent number of standard deviations from the long-term mean for daily rolling mean values. The shaded areas indicate the three drought years analysed - 2010, 2015 and 2020.](display/climvars_anomalies_fullseries_remote.png)

@@

![**Figure S4:** Species sensitivities in drought years 2010, 2015 and 2020. Boxplots represent 25th, 50th (median) and 75th percentile of sensitivity for each species in each year and whiskers represent 95% CIs. All individual sensitivities represented by jittered points.](display/Fig_SI_species_sensitivities.png)

@@

![**Figure S7:** Map of trees with dendrobands in HKK, included in analyses in the 2010, 2015 and 2020 datasets](display/dendro_tree_map.png)

@@

# Alternate model using species random effect only on intercept

![**Figure S17:** Model results from Bayesian causal model of sensitivity as a function of DBH, CII and TWI with species random effect only on intercept.](display/FigSI_coefs_nospre.png)
